---
title: "DSC 630 KSummerville Milestone 3"
author: "Kimberly Summerville"
date: "1/11/2022"
output: word_document
---

```{r}
library(readxl)
library(ggplot2)
library(dplyr)
library(GGally)
library(lubridate)
library(forecast) 
library(caret)
```

Importing datasets
```{r}
State <- data.frame(read.csv("C:/Users/a565448/OneDrive - Fidelity Technology Group, LLC/Personal/School/DSC 630 Predictive Analysis/Final Project/RDC_Inventory_Core_Metrics_State_Monthly_History.csv"))

Mtg <- data.frame(read_excel("C:\\Users\\a565448\\OneDrive - Fidelity Technology Group, LLC\\Personal\\School\\DSC 630 Predictive Analysis\\Final Project\\Monthly-National-Statistics-for-New-Residential-Mortgages-in-the-United-States.xlsx", sheet = "ALL_LOANS", col_names = FALSE, skip = 9))
```


Updating column names
``` {r}
names(Mtg) <- c('Month', 
             'Number of Originations (1,000s)', 
             'Average Loan Amount (1,000 $)', 
             'Average Purchase Price / Appraised Value (1,000 $)',
             'Average Contract Rate', 
             'Percent Share Owner-Occupied Property', 
             'Home Purchase Originations - First Time Homebuyer Loans %', 
             'Home Purchase Originations - Repeat Homebuyer Loans %', 
             'Home Purchase Originations - All Home Purchase Loans %', 
             'Refinance Originations - Cashout Refinance %', 
             'Refinance Originations - Rate and Term Refinance Loans %', 
             'Refinance Originations - All Refinance Loans %',
             'Average Term to Maturity (Yrs.)', 
             'Mortgage Terms - Adjustable-Rate Mortgages %', 
             'Mortgage Terms - 15-Year Fixed Rate Mortgages %', 
             'Mortgage Terms - 20- and 30-Year Fixed Rate Mortgages %', 
             'Back-End Debt-to-Income Ratio - Average DTI', 
             'Back-End Debt-to-Income Ratio - % Less Than or Equal To 36', 
             'Back-End Debt-to-Income Ratio - % More than 36 To 43', 
             'Back-End Debt-to-Income Ratio - % Greater than 43', 
             'All Borrowers Credit Score - Average Credit Score', 
             'All Borrowers Credit Score - % Very Poor Credit', 
             'All Borrowers Credit Score - % Poor Credit', 
             'All Borrowers Credit Score - % Fair Credit', 
             'All Borrowers Credit Score - % Good Credit', 
             'All Borrowers Credit Score - % Excellent Credit',
             'Loan-to-Value - Average Loan-to-Value',
             'Loan-to-Value - Average Combined Loan-to-Value',
             'Loan-to-Value - 70.0% or Less Percent Share by CLTV',
             'Loan-to-Value - 70.1 - 80.0% Percent Share by CLTV',
             'Loan-to-Value - 80.1 - 90.0% Percent Share by CLTV',
             'Loan-to-Value - 90.1 - 95.0% Percent Share by CLTV',
             'Loan-to-Value - 95.1 - 97.0% Percent Share by CLTV',
             'Loan-to-Value - Greater Than 97.0% Percent Share by CLTV',
             'Percent Market Share - Government',
             'Percent Market Share - Enterprise',
             'Percent Market Share - Other Conforming (Portfolio)',
             'Percent Market Share - Jumbo Market',
             'Percent Share of Loans by All Borrowers’ Race - All White alone',
             'Percent Share of Loans by All Borrowers’ Race - All Black or African American alone',
             'Percent Share of Loans by All Borrowers’ Race - All Asian alone',
             'Percent Share of Loans by All Borrowers’ Race - All Native Hawaiian and Other Pacific Islander alone',
             'Percent Share of Loans by All Borrowers’ Race - All American Indian and Alaska Native alone',
             'Percent Share of Loans by All Borrowers’ Race - Multiple races',
             'Percent Share of Loans by All Borrowers’ Ethnicity - All Hispanic or Latino',
             'Percent Share of Loans by All Borrowers’ Ethnicity - At least one Hispanic or Latino',
             'Percent Share of Loans by All Borrowers’ Ethnicity - All White alone, not Hispanic or Latino',
             'Percent Share of Loans by All Borrowers’ Ethnicity - Multiple races or race other than White, not Hispanic or Latino',
             'All Borrowers Ages - Average Age',
             'All Borrowers Ages - % Less Than 25',
             'All Borrowers Ages - % 25 to 34',
             'All Borrowers Ages - % 35 to 44',
             'All Borrowers Ages - % 45 to 54',
             'All Borrowers Ages - % 55 to 64',
             'All Borrowers Ages - % At Least 65',
             'Percent Share of Loans by Number of Borrowers and Gender - Single Borrower - Male',
             'Percent Share of Loans by Number of Borrowers and Gender - Single Borrower - Female',
             'Percent Share of Loans by Number of Borrowers and Gender - Two Borrowers',
             'Percent Share of Loans by Number of Borrowers and Gender - More than Two Borrowers'
)
```

```{r}
Mtg$`Percent Share of Loans by All Borrowers’ Ethnicity - Multiple races or race other than White, not Hispanic or Latino`
```

```{r}
ggplot(Mtg, aes(x = Month)) + 
  geom_line(aes(y = `Refinance Originations - All Refinance Loans %`*3, color = 'blue')) + 
  geom_line(aes(y = `Average Loan Amount (1,000 $)`, color = 'red')) +
  scale_y_continuous() +
  labs(y = "Refinancing % (red) & Average Loan Amount (blue)")
```

```{r}
Mtg_model <- {`Number of Originations (1,000s)` ~ `Average Contract Rate`}

set.seed(1000)

train_and_test <- c(rep(1, trunc((2/3)*nrow(Mtg))), rep(2, trunc((1/3)*nrow(Mtg))))

Mtg$train <- sample(train_and_test)

Mtg$train <- factor(Mtg$train, levels = c(1, 2), labels = c("TRAIN", "TEST"))

Mtg.Train <- subset(Mtg, train == "TRAIN")
Mtg.Test <- subset(Mtg, train == "TEST")

head(Mtg.Train)
head(Mtg.Test)
```

``` {r}
train.model.fit <- lm(Mtg_model, data = Mtg.Train)

Mtg.Train$Predict_Attend <- predict(train.model.fit)

Mtg.Test$Predict_Attend <- predict(train.model.fit, newdata = Mtg.Test)

cat("\n","Proportion of Test Set Variance Accounted for: ", round(cor(Mtg.Test$`Home Purchase Originations - First Time Homebuyer Loans %`, Mtg.Test$Predict_Attend)^2, digits=3), "\n", sep="")

Mtg.Training_Test <- rbind(Mtg.Train, Mtg.Test)
```


``` {r}
ggplot(Mtg.Training_Test, aes(x=`Number of Originations (1,000s)`, y=Predict_Attend)) + 
        geom_line() + 
        facet_wrap(~train) +
        ggtitle("Regression Model Performance : Day of the Week and Attendance") +
        theme(plot.title = element_text(lineheight=3, face="bold", color="black", size=10)) +
        xlab("Actual Originations (Thousands)") +
        ylab("Predicted Originations (Thousands)")

```


```{r}
Mtg_model.fit <- lm(Mtg_model, data = Mtg)  # use all available data
print(summary(Mtg_model.fit))
```
With the model fitted, we can see that as the average borrower age increases by 1 year, the percentage of Originations that are for First Time Homebuyers goes down by 1.987%, which is what we would expect.

``` {r}
Refi_model <- {`Refinance Originations - All Refinance Loans %` ~ `Average Contract Rate`}

set.seed(1000)

train_and_test <- c(rep(1, trunc((2/3)*nrow(Mtg))), rep(2, trunc((1/3)*nrow(Mtg))))

Mtg$train <- sample(train_and_test)

Mtg$train <- factor(Mtg$train, levels = c(1, 2), labels = c("TRAIN", "TEST"))

Mtg.Train <- subset(Mtg, train == "TRAIN")
Mtg.Test <- subset(Mtg, train == "TEST")

train.model.fit <- lm(Mtg_model, data = Mtg.Train)

Mtg.Train$Predict_Attend <- predict(train.model.fit)

Mtg.Test$Predict_Attend <- predict(train.model.fit, newdata = Mtg.Test)

cat("\n","Proportion of Test Set Variance Accounted for: ", round(cor(Mtg.Test$`Home Purchase Originations - First Time Homebuyer Loans %`, Mtg.Test$Predict_Attend)^2, digits=3), "\n", sep="")

Mtg.Training_Test <- rbind(Mtg.Train, Mtg.Test)

ggplot(Mtg.Training_Test, aes(x=`Refinance Originations - All Refinance Loans %`, y=Predict_Attend)) + 
        geom_line() + 
        facet_wrap(~train) +
        ggtitle("Regression Model Performance : % of Originations being Refinance Loans") +
        theme(plot.title = element_text(lineheight=3, face="bold", color="black", size=10)) +
        xlab("Actual Originations (Thousands)") +
        ylab("Predicted Originations (Thousands)")
```

```{r}
Mtg_model.fit <- lm(Mtg_model, data = Mtg)  # use all available data
print(summary(Mtg_model.fit))
```
Here we see that for each point that the Average Contract Rate (Interest Rate) goes up, the percentage of Originations being from Refinances goes down 2.477 points, which is expected.

```{r}
model <- train(`Number of Originations (1,000s)` ~ `Average Contract Rate`, data = Mtg, method = "lm")
model
```


```{r}
Orig <- Mtg$`Number of Originations (1,000s)`

mtg <- ts(Orig, start = c(1998, 1), end = c(2021, 3), 
                            deltat = 1/12)

fit <- auto.arima(mtg)

plot(forecast(fit, 25), xlab = "Month", ylab = "Number of Originations", main = "Originations Predictions", col.main ="darkgreen")
```

```{r}
forecast(fit, 25)
```
```{r}
Int_model <- {`Number of Originations (1,000s)` ~ `Average Contract Rate`}

Int_model.fit <- lm(Int_model, data = Mtg)  # use all available data
print(summary(Int_model.fit))
```

```{r}
IntRefi_model <- {`Refinance Originations - All Refinance Loans %` ~ `Average Contract Rate`}

IntRefi_model.fit <- lm(IntRefi_model, data = Mtg)  # use all available data
print(summary(IntRefi_model.fit))
```

```{r}
IntPurch_model <- {`Home Purchase Originations - All Home Purchase Loans %` ~ `Average Contract Rate`}

IntPurch_model.fit <- lm(IntPurch_model, data = Mtg)  # use all available data
print(summary(IntPurch_model.fit))


```

```{r}
ggpairs(data=Mtg, columns=2:5, title="Interest")
fit_ex <- lm(`Number of Originations (1,000s)` ~ `Average Contract Rate`, data = Mtg)
ggplot(data=Mtg, aes(fit_ex$residuals)) +
geom_histogram(binwidth = 1, color = "green", fill = "yellow") +
theme(panel.background = element_rect(fill = "red"),
axis.line.x=element_line(),
axis.line.y=element_line()) +
ggtitle("Histogram for women height")
ggplot(data = Mtg, aes(x = `Number of Originations (1,000s)`, y = `Average Contract Rate`)) +
geom_point() +
stat_smooth(method = "lm", col = "blue") +
theme(panel.background = element_rect(fill = "grey"),
axis.line.x=element_line(),
axis.line.y=element_line()) +
ggtitle("Linear Model fitted to above data")
predict(fit_ex, data.frame(weight = 70.2))
```

```{r}
Greater <- Mtg %>% 
  filter(`Average Contract Rate` >= 5.50)

Less <- Mtg %>% 
  filter(`Average Contract Rate` <= 5.50)

head(Greater)
head(Less)
```


```{r} 
#Greater than 5.5% interest rate model
ggpairs(data=Greater, columns=2:5, title="Interest")
fit_ex <- lm(`Number of Originations (1,000s)` ~ `Average Contract Rate`, data = Greater)
ggplot(data=Greater, aes(fit_ex$residuals)) +
geom_histogram(binwidth = 1, color = "green", fill = "yellow") +
theme(panel.background = element_rect(fill = "red"),
axis.line.x=element_line(),
axis.line.y=element_line()) +
ggtitle("Histogram for women height")
ggplot(data = Greater, aes(x = `Number of Originations (1,000s)`, y = `Average Contract Rate`)) +
geom_point() +
stat_smooth(method = "lm", col = "blue") +
theme(panel.background = element_rect(fill = "grey"),
axis.line.x=element_line(),
axis.line.y=element_line()) +
ggtitle("Interest Rate Greater than 5.5%")
predict(fit_ex, data.frame(weight = 70.2))
```

```{r} 
#Lesser than 5.5% interest rate model
ggpairs(data=Less, columns=2:5, title="Interest")
fit_ex <- lm(`Number of Originations (1,000s)` ~ `Average Contract Rate`, data = Less)
ggplot(data=Less, aes(fit_ex$residuals)) +
geom_histogram(binwidth = 1, color = "green", fill = "yellow") +
theme(panel.background = element_rect(fill = "red"),
axis.line.x=element_line(),
axis.line.y=element_line()) +
ggtitle("Histogram for women height")
ggplot(data = Less, aes(x = `Number of Originations (1,000s)`, y = `Average Contract Rate`)) +
geom_point() +
stat_smooth(method = "lm", col = "blue") +
theme(panel.background = element_rect(fill = "grey"),
axis.line.x=element_line(),
axis.line.y=element_line()) +
ggtitle("Interest Rate Less than 5.5%")
predict(fit_ex, data.frame(weight = 70.2))
```



```{r}
Mtg2009 <- Mtg %>% 
  dplyr::filter(Month >= as.POSIXct("2008-12-01"))

Mtg1998 <- Mtg %>% 
  dplyr::filter(Month <= as.POSIXct("2008-12-01"))
```

```{r}
Orig <- Mtg2009$`Number of Originations (1,000s)`

mtg <- ts(Orig, start = c(2009, 1), end = c(2021, 3), 
                            deltat = 1/12)

fit <- auto.arima(mtg)

plot(forecast(fit, 25), xlab = "Month", ylab = "Number of Originations", main = "Originations Predictions", col.main ="darkgreen")

forecast(fit, 25)
```

```{r}
Orig <- Mtg1998$`Number of Originations (1,000s)`

mtg <- ts(Orig, start = c(1998, 1), end = c(2009, 1), 
                            deltat = 1/12)

fit <- auto.arima(mtg)

plot(forecast(fit, 25), xlab = "Month", ylab = "Number of Originations", main = "Originations Predictions", col.main ="darkgreen")

forecast(fit, 25)
```
















